### 2020-07-03 paper/paper.rmd
Adding a JOSS paper ref #25

---
title: 'oolong: An R package for validating  automated content analysis tools'
tags:
  - R
  - text analysis
  - topic model
  - sentiment analysis
  - validation
authors:
  - name: Chung-hong Chan
    orcid: 0000-0003-0872-7098
    affiliation: 1
  - name: Marius Sältzer
    orcid: 0000-0002-8604-4666
    affiliation: 1
affiliations:
 - name: Mannheimer Zentrum für Europäische Sozialforschung, Universität Mannheim
   index: 1
citation_author: Chan & Sältzer.
date: 3 July 2020
year: 2020
bibliography: paper.bib
output: rticles::joss_article
csl: apa.csl
journal: JOSS
---

# Summary

Oolong is an R package providing functions for validating common automated content analysis techniques such as topic modeling and dictionary-based methods. This package is designed for R users needing to validate these methods. Typical users of oolong are communication and political scientists, among others.

Validity is a requirement of content analysis [@krippendorff2018content; @neuendorf2016content]. Validation of automated methods has been called for by many scholars, e.g. @grimmer2013text; @ribeiro2016sentibench; @van2018communication. However, the method for validating these automated methods is unstandardized.

Oolong makes it easy to generate standard validation tests suggested by @chang2009reading and @song2020validations.

# Validating topic models

Topic models can be validated by word intrusion test and topic intrusion test [@chang2009reading]. In these tests, a human rater is asked to pick an odd word from a bunch of words (word intrusion test) or pick an odd topic from a bunch of topics for a document (topic intrusion test). Oolong provides an easy-to-use Shiny interface for these tests (Figure 1).

```{r, echo = FALSE, out.width="50%", fig.cap = "A screenshot of word intrusion test"}
knitr::include_graphics("paper_files/fig1.png")
```

Currently, oolong supports a variety of topic models, e.g. structural topic models / correlated topic models from `stm` [@roberts2019stm], warp-LDA models from `text2vec` [@selivanov2020tex2vec], latent dirichlet allocation / correlated-topic models from `topicmodels` [@bettina2011topicmodels], biterm topic models from `BTM` [@wijffels2020btm] and keyword-assisted topic models from `keyATM` [@eshima2020keyatm].

For instance, `abstracts_stm` is a structural topic model trained with the text data from `abstracts$text` [@chan2020high].

```{r, message = FALSE}
library(stm)
library(tibble)
library(dplyr)
library(quanteda)
library(oolong)
```

```{r}
abstracts_stm
```

```{r}
abstracts
```

The function `create_oolong` creates a test object with both word intrusion test and topic intrusion test.

```{r}
oolong_test <- create_oolong(input_model = abstracts_stm,
                             input_corpus = abstracts$text)
oolong_test
```

The tests can be administered with methods `do_word_intrusion_test` and `do_topic_intrusion_test`.

```r
oolong_test$do_word_intrusion_test()
oolong_test$do_topic_intrusion_test()
```

After both tests has been done by a human rater, the test object must be locked and then accuracy metrics such as model precision (MP) and TLO (topic log odd) are displayed.

```{r, include = FALSE}
genius_topic <- function(obj1) {
    obj1$.__enclos_env__$private$test_content$topic$answer <- obj1$.__enclos_env__$private$test_content$topic$intruder
    return(obj1)
}
genius_word <- function(obj1) {
    obj1$.__enclos_env__$private$test_content$word$answer <- obj1$.__enclos_env__$private$test_content$word$intruder
    return(obj1)
}
oolong_test <- genius_word(genius_topic(oolong_test))
oolong_test$.__enclos_env__$private$test_content$word$answer[1] <- "wronganswer"

oolong_test$.__enclos_env__$private$test_content$topic$answer[2] <- sample(oolong_test$.__enclos_env__$private$test_content$topic$candidates[[2]], 1)
```

```{r lock}
oolong_test$lock()
oolong_test
```

The suggested workflow is to have at least two human raters to do the same set of tests. Test object can be cloned to allow multiple raters to do the test. More than one test object can be studied together using the function `summarize_oolong()`.

```{r}
oolong_test_rater1 <- create_oolong(abstracts_stm, abstracts$text)
oolong_test_rater2 <- clone_oolong(oolong_test_rater1)
```

```r
## Let rater 1 do the test.
oolong_test_rater1$do_word_intrusion_test()
oolong_test_rater1$do_topic_intrusion_test()
oolong_test_rater1$lock()

## Let rater 2 do the test.
oolong_test_rater2$do_word_intrusion_test()
oolong_test_rater2$do_topic_intrusion_test()
oolong_test_rater2$lock()
```

```{r, include = FALSE}
### Mock this process
set.seed(46709394)
oolong_test_rater1 <- oolong:::.monkey_test(oolong_test_rater1, intelligent = 0.3)
oolong_test_rater2 <- oolong:::.monkey_test(oolong_test_rater2, intelligent = 0)
oolong_test_rater1$lock()
oolong_test_rater2$lock()
```

Get a summary of the two objects.

```{r, step3}
summarize_oolong(oolong_test_rater1, oolong_test_rater2)
```

# Validating dictionary-based methods

Dictionary-based methods such as AFINN [@nielsen2011new] can be validated by creating a gold standard dataset [@song2020validations]. Oolong provides a workflow for generating such gold standard dataset.

For example, you are interested in studying the sentiment of tweets from Donald Trump. `trump2k` is a random subset of 2,000 tweets from Donald Trump. And you would like to use AFINN to extract sentiment from these tweets. In this analysis, AFINN sentiment score is the *target value*.

```{r}
tibble(text = trump2k)
```

A test object can be generated also with `create_oolong`. The argument `construct` should be an adjective, e.g. "positive" or "liberal".

```{r}
trump <- create_oolong(input_corpus = trump2k,
                       construct = "positive",
                       exact_n = 20)
trump
```

Similarly, we suggest to have at least two human coders to do the same set of tests.

```{r}
trump2 <- clone_oolong(trump)
```

Instruct two coders to code the tweets and lock the objects.

```r
trump$do_gold_standard_test()
trump2$do_gold_standard_test()
trump$lock()
trump2$lock()
```

```{r, include = FALSE}
trump$.__enclos_env__$private$test_content$gold_standard <- 
structure(list(case = 1:20, text = c("Thank you Eau Claire, Wisconsin. \n#VoteTrump on Tuesday, April 5th!\nMAKE AMERICA GREAT AGAIN! https://t.co/JI5JqwHnMC", 
"\"@bobby990r_1: @realDonaldTrump would lead polls the second he announces candidacy! America is waiting for him to LEAD us out of this mess!", 
"\"@KdanielsK: @misstcassidy @AllAboutTheTea_ @realDonaldTrump My money is on Kenya getting fired first.\"", 
"Thank you for a great afternoon Birmingham, Alabama! #Trump2016 #MakeAmericaGreatAgain https://t.co/FrOkqCzBoD", 
"\"@THETAINTEDT: @foxandfriends @realDonaldTrump Trump 2016 http://t.co/UlQWGKUrCJ\"", 
"People believe CNN these days almost as little as they believe Hillary....that's really saying something!", 
"It was great being in Michigan. Remember, I am the only presidential candidate who will bring jobs back to the U.S.and protect car industry!", 
"\"@DomineekSmith: @realDonaldTrump is the best Republican presidential candidate of all time.\"  Thank you.", 
"Word is that little Morty Zuckerman’s @NYDailyNews loses more than $50 million per year---can that be possible?", 
"\"@Chevy_Mama: @realDonaldTrump I'm obsessed with @celebrityapprenticeNBC. Honestly,  Mr Trump, you are very inspiring\"", 
"President Obama said \"ISIL continues to shrink\" in an interview just hours before the horrible attack in Paris. He is just so bad! CHANGE.", 
".@HillaryClinton loves to lie. America has had enough of the CLINTON'S! It is time to #DrainTheSwamp! Debates https://t.co/3Mz4T7qTTR", 
"\"@jerrimoore: @realDonaldTrump awesome. A treat to get to see the brilliant Joan Rivers once more #icon\"", 
"\"@shoegoddesss: @realDonaldTrump Will definitely vote for you. Breath of fresh air. America needs you!\"", 
"Ted is the ultimate hypocrite. Says one thing for money, does another for votes. \nhttps://t.co/hxdfy0mjVw", 
"\"@Lisa_Milicaj: Truth be told, I  never heard of The National Review until they \"tried\" to declare war on you. No worries, you got my vote!\"", 
"THANK YOU Daytona Beach, Florida!\n#MakeAmericaGreatAgain https://t.co/IAcLfXe463", 
"People rarely say that many conservatives didn’t vote for Mitt Romney. If I can get them to vote for me, we win in a landslide.", 
"Trump National Golf Club, Washington, D.C. is on 600 beautiful acres fronting the Potomac River. A fantastic setting! http://t.co/pYtkbyKwt5", 
"\"@DRUDGE_REPORT: REUTERS 5-DAY ROLLING POLL: TRUMP 34%, CARSON 19.6%, RUBIO 9.7%, CRUZ 7.7%...\" Thank you - a great honor!"
), answer = c(4L, 4L, 2L, 5L, 3L, 2L, 4L, 5L, 2L, 4L, 1L, 1L, 
4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L), target_value = c(NA, NA, NA, 
NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 
NA)), row.names = c(NA, -20L), class = c("tbl_df", "tbl", "data.frame"
                                         ))

trump2$.__enclos_env__$private$test_content$gold_standard <- 
structure(list(case = 1:20, text = c("Thank you Eau Claire, Wisconsin. \n#VoteTrump on Tuesday, April 5th!\nMAKE AMERICA GREAT AGAIN! https://t.co/JI5JqwHnMC", 
"\"@bobby990r_1: @realDonaldTrump would lead polls the second he announces candidacy! America is waiting for him to LEAD us out of this mess!", 
"\"@KdanielsK: @misstcassidy @AllAboutTheTea_ @realDonaldTrump My money is on Kenya getting fired first.\"", 
"Thank you for a great afternoon Birmingham, Alabama! #Trump2016 #MakeAmericaGreatAgain https://t.co/FrOkqCzBoD", 
"\"@THETAINTEDT: @foxandfriends @realDonaldTrump Trump 2016 http://t.co/UlQWGKUrCJ\"", 
"People believe CNN these days almost as little as they believe Hillary....that's really saying something!", 
"It was great being in Michigan. Remember, I am the only presidential candidate who will bring jobs back to the U.S.and protect car industry!", 
"\"@DomineekSmith: @realDonaldTrump is the best Republican presidential candidate of all time.\"  Thank you.", 
"Word is that little Morty Zuckerman’s @NYDailyNews loses more than $50 million per year---can that be possible?", 
"\"@Chevy_Mama: @realDonaldTrump I'm obsessed with @celebrityapprenticeNBC. Honestly,  Mr Trump, you are very inspiring\"", 
"President Obama said \"ISIL continues to shrink\" in an interview just hours before the horrible attack in Paris. He is just so bad! CHANGE.", 
".@HillaryClinton loves to lie. America has had enough of the CLINTON'S! It is time to #DrainTheSwamp! Debates https://t.co/3Mz4T7qTTR", 
"\"@jerrimoore: @realDonaldTrump awesome. A treat to get to see the brilliant Joan Rivers once more #icon\"", 
"\"@shoegoddesss: @realDonaldTrump Will definitely vote for you. Breath of fresh air. America needs you!\"", 
"Ted is the ultimate hypocrite. Says one thing for money, does another for votes. \nhttps://t.co/hxdfy0mjVw", 
"\"@Lisa_Milicaj: Truth be told, I  never heard of The National Review until they \"tried\" to declare war on you. No worries, you got my vote!\"", 
"THANK YOU Daytona Beach, Florida!\n#MakeAmericaGreatAgain https://t.co/IAcLfXe463", 
"People rarely say that many conservatives didn’t vote for Mitt Romney. If I can get them to vote for me, we win in a landslide.", 
"Trump National Golf Club, Washington, D.C. is on 600 beautiful acres fronting the Potomac River. A fantastic setting! http://t.co/pYtkbyKwt5", 
"\"@DRUDGE_REPORT: REUTERS 5-DAY ROLLING POLL: TRUMP 34%, CARSON 19.6%, RUBIO 9.7%, CRUZ 7.7%...\" Thank you - a great honor!"
), answer = c(5L, 3L, 2L, 5L, 3L, 1L, 4L, 5L, 2L, 4L, 1L, 1L, 
4L, 4L, 2L, 4L, 4L, 4L, 4L, 4L), target_value = c(NA, NA, NA, 
NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, NA, 
NA)), row.names = c(NA, -20L), class = c("tbl_df", "tbl", "data.frame"
                                         ))
trump$lock()
trump2$lock()
```

The method `turn_gold` converts a test object into a quanteda corpus [@benoit2018quanteda].

```{r}
gold_standard <- trump$turn_gold()
gold_standard
```

This corpus can be used to calculate the target value, e.g. AFINN.

```{r}
dfm(gold_standard, remove_punct = TRUE) %>% dfm_lookup(afinn) %>%
    quanteda::convert(to = "data.frame") %>%
    mutate(matching_word_valence = (neg5 * -5) + (neg4 * -4) +
               (neg3 * -3) + (neg2 * -2) + (neg1 * -1) +
               (zero * 0) + (pos1 * 1) + (pos2 * 2) + (pos3 * 3) +
               (pos4 * 4) + (pos5 * 5),
           base = ntoken(gold_standard, remove_punct = TRUE),
           afinn_score = matching_word_valence / base) %>% 
		   pull(afinn_score) -> afinn_score
```

Summarize all oolong objects with the target value.

```{r, message = FALSE}
res <- summarize_oolong(trump, trump2, target_value = afinn_score)
```

Printing the summary shows Krippendorff's Alpha, an indicator of interrater reliability. The validity metrics of a text analytic method can be tinted by poor interrater reliability of manual annotations [@song2020validations]. It is important to ensure high interrater reliability first.

```{r}
res
```

Additional diagnostic plots can also be displayed.

```{r diagplot, fig.cap = "Diagnostic plots generated by oolong"}
plot(res)
```

The 4 subplots from left to right, top to bottom are: 1) correlation between human judgement and target value; 2) Bland-Altman plot; 3) correlation between target value and content length and 4) Cook's distance of all data point. These plots are helpful to determine criterion validity, agreement, robustness against content length and outliers of the target value.

# Acknowledgements

The development of oolong is partially supported by SAGE Concept Grant.

# References