### 2021-05-22 deploy.Rmd
adding tests

In oolong 0.3.22, functions for deploying oolong tests were added (`export_oolong`, `revert_oolong` etc.). These functions make it possible for the coders to conduct validation tests online using their browser, rather than having to install R on their computer.

The basic workflow is simple: 1) create an oolong test object as usual; 2) deploy the test online and obtain the URL to the test; 3) ask your coders to conduct the test online and send back the data file; 4) revert back the data file to an oolong object.