### 2024-06-17 methodshub.qmd
MH Example

---
title: oolong - Create Validation Tests for Automated Content Analysis
format:
  html:
    embed-resources: true
  gfm: default
---

## Description

<!-- - Provide a brief and clear description of the method, its purpose, and what it aims to achieve. Add a link to a related paper from social science domain and show how your method can be applied to solve that research question.   -->

Intended to create standard human-in-the-loop validity tests for typical automated content analysis such as topic modeling and dictionary-based methods. This package offers a standard workflow with functions to prepare, administer and evaluate a human-in-the-loop validity test. This package provides functions for validating topic models using word intrusion, topic intrusion (Chang et al. 2009,  <https://papers.nips.cc/paper/3700-reading-tea-leaves-how-humans-interpret-topic-models>) and word set intrusion (Ying et al. 2021) [<doi:10.1017/pan.2021.33>](https://doi.org/10.1017/pan.2021.33) tests. This package also provides functions for generating gold-standard data which are useful for validating dictionary-based methods. The default settings of all generated tests match those suggested in Chang et al. (2009) and Song et al. (2020) [<doi:10.1080/10584609.2020.1723752>](https://doi.org/10.1080/10584609.2020.1723752).

## Keywords

* Validity
* Text Analysis
* Topic Model

## Science Usecase(s)

<!-- - Include usecases from social sciences that would make this method applicable in a certain scenario.  -->
<!-- The use cases or research questions mentioned should arise from the latest social science literature cited in the description. -->

This package was used in the literature to valid topic models and prediction models trained on text data, e.g. [Rauchfleisch et al. (2023)](10.1080/17512786.2022.2110928), [Rothut, et al. (2023)](https://doi.org/10.1177/14614448231164409), [Eisele, et al. (2023)](https://doi.org/10.1080/19312458.2023.2230560).

## Repository structure

This repository follows [the standard structure of an R package](https://cran.r-project.org/doc/FAQ/R-exts.html#Package-structure).

## Environment Setup

With R installed:

```r
install.packages("oolong")
```

<!-- ## Hardware Requirements (Optional) -->
<!-- - The hardware requirements may be needed in specific cases when a method is known to require more memory/compute power.  -->
<!-- - The method need to be executed on a specific architecture (GPUs, Hadoop cluster etc.) -->

## Input Data

<!-- - The input data has to be a Digital Behavioral Data (DBD) Dataset -->
<!-- - You can provide link to a public DBD dataset. GESIS DBD datasets (https://www.gesis.org/en/institute/digital-behavioral-data) -->

The input data has to be a topic model or prediction model trained on text data. For example, one can train a topic model from the text data (tweets from Donald trump) included in the package by:

```r
library(seededlda)
library(quanteda)
trump_corpus <- corpus(trump2k)
tokens(trump_corpus, remove_punct = TRUE, remove_numbers = TRUE, remove_symbols = TRUE,
       split_hyphens = TRUE, remove_url = TRUE) %>% tokens_tolower() %>%
    tokens_remove(stopwords("en")) %>% tokens_remove("@*")  -> trump_toks

model <- textmodel_lda(x = dfm(trump_toks), k = 8, verbose = TRUE)
```

## Sample Input and Output Data

<!-- - Show how the input data looks like through few sample instances -->
<!-- - Providing a sample output on the sample input to help cross check  -->

A sample input is a model trained on text data, e.g.

```{r}
#| message: false
library(oolong)
library(seededlda)
abstracts_seededlda
```

The sample output is an oolong S3 object.

## How to Use

Please refer to the [overview of this package](https://gesistsa.github.io/oolong/articles/overview.html) for a comprehensive introduction of all test types.

Suppose there is a topic model trained on some text data called `abstracts_seededlda`, which is included in the package.

```{r}
library(oolong)
abstracts_seededlda
```

Suppose one would like to conduct a word intrusion test (Chang et al. 2009) to validate this topic model. This test can be generated by the `wi()` function.

```{r}
oolong_test <- wi(abstracts_seededlda, userid = "Hadley")
oolong_test
```

One can then conduct the test following the instruction displayed, i.e. `oolong_test$$do_word_intrusion_test()`.

```r
oolong_test$do_word_intrusion_test()
```

One should see a graphic interface like the following and conduct the test.

<img src="man/figures/oolong_demo.gif" align="center" height="400" />

After the test, one can finalize the test by locking the test.

```{r}
#| include: false
### Mock this process
oolong_test$.__enclos_env__$private$test_content$wi$answer <- oolong_test$.__enclos_env__$private$test_content$wi$intruder
oolong_test$.__enclos_env__$private$test_content$wi$answer[1] <- "wronganswer"
```

```{r}
oolong_test$lock()
```

And then obtain the result of the test. For example:

```{r}
oolong_test
```

## Contact Details

Maintainer: Chung-hong Chan <chainsawtiney@gmail.com>

Issue Tracker: [https://github.com/gesistsa/oolong/issues](https://github.com/gesistsa/oolong/issues)

## Publication

1. Chan, C. H., & Sältzer, M. (2020). oolong: An R package for validating automated content analysis tools. The Journal of Open Source Software: JOSS, 5(55), 2461. https:://doi.org/10.21105/joss.02461

<!-- ## Acknowledgements -->
<!-- - Acknowledgements if any -->

<!-- ## Disclaimer -->
<!-- - Add any disclaimers, legal notices, or usage restrictions for the method, if necessary. -->